Pushdown Control-Flow Analysis via Refunctionalization (Functional Pearl)

### 2018-03-06 paper.tex
reorganize

  In this paper, we show their correspondence and how to syntactically transform small-step 
  abstracting abstract machines into big-step abstracting definitional 
  interpreters.
  The transformations include fusing, disentangling, refunctionalizing 
  and monadification (or un-CPS to direct-style). All of them properly handle the 
  collecting semantics and the non-determinism of abstract interpretation. 
  After each transformation, we also obtain an intermediary form of 
  abstract interpreter.
  Remarkably, refunctionalization sequentializes the abstract evaluation.
  And follow the idea that evaluation contexts are defunctionalized 
  continuations, we reveal how precise call/return match in control-flow 
  analysis can be obtained by refunctionalizing a small-step abstracting 
  abstract machine,
  as well as explain how it been lost by defunctionalizing the higher-order
  continuations into an firts-order data type.

@@

\item We present an abstract interpreter obtained from refunctionalizing
  an small-step abstracting abstract machine. 
  This abstract interpreter naturally has pushdown control-flow property
  since we represent the evaluation contexts of analyzed language using
  continuations of the meta-language, and therefore sequentialize the
  analyzing.
  We also analyze its termination, soundness and complexity.

@@

\begin{tikzpicture}
  \matrix (m) [matrix of math nodes, row sep=3em, column sep=3.7em]
    { 
      \begin{smallmatrix} \text{abstracting} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
      \begin{smallmatrix} \text{big-step} \\ \text{abstracting} \\ \text{abstract} \\ \text{machines} \end{smallmatrix}  & 
      \begin{smallmatrix} \text{disentangled}  \\ \text{abstracting} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
      \begin{smallmatrix} \text{refunctionalized}  \\ \text{abstracting} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
      \begin{smallmatrix} \text{abstracting} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix}  \\
      \begin{smallmatrix} \text{abstract} \\ \text{machines} \end{smallmatrix} & 
        &
      %\begin{smallmatrix} \text{big-step} \\ \text{abstract} \\ \text{machines} \end{smallmatrix}  & 
        &
      %\begin{smallmatrix} \text{disentangled} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
        &
      %\begin{smallmatrix} \text{refunctionalized} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
      \begin{smallmatrix} \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\ 
    };

  { 
    [start chain] \chainin (m-1-1);
    \chainin (m-1-2) [join={node[above,labeled] {\text{fusion}}}];

    \chainin (m-1-3) [join={node[above,labeled] {\text{disentangling}}}];

    \chainin (m-1-4) [join={node[above,labeled] {\text{refunc.}}}];

    \chainin (m-1-5) [join={node[above,labeled] {\text{monadify}}}];
  }
  { 
    [start chain] 
    \chainin (m-2-1);
    { [start branch=A] \chainin (m-1-1)
      [join={node[left, labeled] {\alpha}}];
    }
  }
    %\chainin (m-2-2);
    %\chainin (m-2-3);
    %\chainin (m-2-4);
  {
    [start chain] 
    \chainin (m-2-5); 
    { [start branch=B] \chainin (m-1-5)
      [join={node[left, labeled] {\alpha}}];
    }
  }
\end{tikzpicture}

@@

2) the precise call/return match problem in control-flow analysis and its existing solutions.

@@

@@

\section{From AAM to ADI}

\subsection{Fusing}

\subsection{Disentangling}

%probably Store Widening \& Abstract Garbage Collection
%monadic
\subsection{Monadification: to Abstracting Definitional Interpreter}

@@

\subsection{Un-CPS: to Yet Another Abstracting Definitional Interpreter}

@@

@@

\section{Detour: Shiver's \textit{k}-CFA?}

>>>>

  In this paper, we show their correspondence and how to syntactically transform small-step 
  abstracting abstract machines into big-step abstracting definitional 
  interpreters.
  The transformations include fusing, disentangling, refunctionalizing 
  and monadification (or un-CPS to direct-style with delimited continuation). 
  All of them properly handle the 
  collecting semantics and the non-determinism of abstract interpretation. 
  After each transformation, we also obtain an intermediary form of 
  abstract interpreter.
  Remarkably, refunctionalization sequentializes the abstract evaluation.
  And follow the idea that evaluation contexts are defunctionalized 
  continuations, we reveal how precise call/return match in control-flow 
  analysis can be obtained by refunctionalizing a small-step abstracting 
  abstract machine,
  as well as explain how it been lost by defunctionalizing the higher-order
  continuations into an firts-order data type.

@@

\item A new understanding of pushdown control-flow analysis by 
  refunctionalization.
  We present an abstract interpreter obtained from refunctionalizing
  an small-step abstracting abstract machine. 
  This abstract interpreter naturally has pushdown control-flow property
  because the refunctionalization forces us to sequentialize the 
  abstract exectuion.
  We also analyze its termination and soundness.

@@

\begin{tikzpicture}
  \matrix (m) [matrix of math nodes, row sep=2.5em, column sep=3.7em]
    { 
        &  &  &  & 
      \begin{smallmatrix} \text{abstracting} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\
      \begin{smallmatrix} \text{abstracting} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
      \begin{smallmatrix} \text{big-step} \\ \text{abstracting} \\ \text{abstract} \\ \text{machines} \end{smallmatrix}  & 
      \begin{smallmatrix} \text{disentangled}  \\ \text{abstracting} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
      \begin{smallmatrix} \text{refunctionalized}  \\ \text{abstracting} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
      \begin{smallmatrix} \text{monadic} \\ \text{abstracting} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix}  \\
      \begin{smallmatrix} \text{abstract} \\ \text{machines} \end{smallmatrix} & 
        &
      %\begin{smallmatrix} \text{big-step} \\ \text{abstract} \\ \text{machines} \end{smallmatrix}  & 
        &
      %\begin{smallmatrix} \text{disentangled} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
        &
      %\begin{smallmatrix} \text{refunctionalized} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
      \begin{smallmatrix} \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\ 
    };
  {
    [start chain] \chainin (m-2-4);
    \chainin (m-1-5) [join={node[sloped,anchor=center,above,labeled] {\text{un-CPS}}}];
    \chainin (m-2-5) [join={node[above,labeled] {\text{}}}];
    \chainin (m-1-5) [join={node[above,labeled] {\text{}}}];
  }
  { 
    [start chain] \chainin (m-2-1);
    \chainin (m-2-2) [join={node[above,labeled] {\text{fusion}}}];
    \chainin (m-2-3) [join={node[above,labeled] {\text{disentangling}}}];
    \chainin (m-2-4) [join={node[above,labeled] {\text{refunc.}}}];
    \chainin (m-2-5) [join={node[above,labeled] {\text{monadify}}}];
  }
  { 
    [start chain] 
    \chainin (m-3-1);
    { [start branch=A] \chainin (m-2-1)
      [join={node[left, labeled] {\alpha}}];
    }
  }
    %\chainin (m-2-2);
    %\chainin (m-2-3);
    %\chainin (m-2-4);
  {
    [start chain] 
    \chainin (m-3-5); 
    { [start branch=B] \chainin (m-2-5)
      [join={node[right, labeled] {\alpha}}];
    }
  }
\end{tikzpicture}

@@

@@

\section{Fusing}

\section{Disentangling}

\section{Refunctionalization}

@@

%probably Store Widening \& Abstract Garbage Collection
%monadic
\section{Monadification: to Abstracting Definitional Interpreter}

@@

\section{Un-CPS: to Yet Another Abstracting Definitional Interpreter}

@@

May be insert to section refunc.

@@

%\section{Detour: Shiver's \textit{k}-CFA?}
